---
title: "notes"
author: "Luka Loncar"
date: '27 listopada 2017 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Import 

Making a single data frame out of many data frames
Creating features that make sense

# Data Understanding and Preparation

Are there any NA's (+), missclasified data types(+), errors and outliners?
We achieve that through descriptive and exploratory statistics. Each is complement to one another. 

- DS

- EDA by data type

```{r}
glimpse(train)

table(train$pickup_neighborhood)
```
TO DO: trip_duration, trip_distance,fare_amount, tolls_amount TMIN TMAX


passenger_count - what is when a passanger is 0? haha 
trip_distance - 7 outliners
extra - one outliner
pickup_neighborhood - make classes with less than 1000 observations into one class. 
dropoff_borough - Staten Island have only 6 observations
dropoff_neighborhood - same as pickup_neighborhood


```{r}
glimpse(taxi_data)
```


```{r}
library(dplyr)
strigar <- data.frame(ifelse(taxi_data$passenger_count < 3, 1, taxi_data$passenger_count))
View(strigar)
strigar$column <- strigar$ifelse.taxi_data.passenger_count...3..1..taxi_data.passenger_count.

strigar[,1] <- NULL 



# %in% is a logical comparision


?ifelse
n_distinct(taxi_data$passenger_count)

any(taxi_data$passenger_count == 0)

taxi_dataa <- taxi_data

taxi_dataa$passenger_count <- ifelse(taxi_dataa$passenger_count < 4, 1,taxi_dataa$passenger_count)

any(taxi_dataa$passenger_count == 6)
n_distinct(taxi_dataa$passenger_count)

summarize(count = n(taxi_dataa$passenger_count))

summarize(taxi_dataa, count = n())
          
          
taxi_dataa %>%
  group_by(passenger_count) %>%
  summarize(n())
          
yes <- 1:3
no <- pi^(0:3)

typeof(ifelse(NA,    yes, no))
typeof(ifelse(TRUE,  yes, no))
typeof(ifelse(FALSE, yes, no))
```







