---
title: "Data Understandind & Preparation"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)
```

## Reading the data

* Import the data

```{r}

sample <- read.csv("sample100000.csv")

```

### What does each column mean?

```{r echo=FALSE}

str(sample)

```

* **Sample** has 100.000 observations and 13 variables
* **pickup_datetime** and **dropoff_datetime** are date & time information for each trip. The variable type is *factor* for both variables.
* **borough** and **neihgborhood** level pick up and dropoff locations. Both variables are factor type.
* **passenger_count** is the number of passengers for each trip.integer type.
* **payment_type** A numeric code signifying how the passenger paid for the trip.variable type is integer.  
1= Credit card  
2= Cash  
3= No charge  
4= Dispute  
* **trip_distance** in miles.
* **fare_amount** The time-and-distance fare calculated.
* **extra** Miscellaneous extras and surcharges.(Currently, this only includes
the $0.50 and $1 rush hour and overnight charges.)
* **tip_amount** This field is automatically populated for credit card
tips. (Cash tips are not included.)
* **tolls_amount** Total amount of all tolls paid in trip.  
  
### How can we make our dataset more "clean"?